Discovering Focused Microgenre Communities

### 2023-08-04 main.tex
Updates from Overleaf

\subsection{Precursors of the Proposed Approach}
The approach proposed here builds on the core ideas of ``duality'' and ``relationality'' that have become foundational for formal approaches to measuring culture in the sociology of taste \citep{mutzel2020duality, mohr2015formal}. In this sense, I followed a path already trailed by others. For instance, Goldberg's \citeyearpar{goldberg2011mapping} {\em Relational Class Analysis} (RCA) ``classifies the people'' (in the terms developed below when reviewing standard data-analytic approaches in the sociology of taste) included a survey by partitioning a person-by-person matrix built from higher order similarities (sums of differences of differences) between each pair of person's corresponding response vector coding their stances (e.g., like, dislike, neutral) toward a set of cultural objects (e.g., genres, attitude items). Boutyline's 
\citeyearpar{boutyline2017improving} {\em Correlational Class Analysis} (CCA) also classifies the people but this time by using the absolute value of the correlation distance between response vectors to build the person-by-person similarity matrix. Boutyline and Vaisey's {\em Belief Network Analysis} (BNA) classifies the objects people connect to (in their case, beliefs, but could also be genres) by creating a weighted object network, where the weight of the tie between two objects (attitude toward government spending and attitude toward immigration) is given by the magnitude of their correlation coefficient in the survey.

As we will see, I propose a complementary and synergistic approach with these previous efforts. More to the point, the relational identification of microgenres should be able to improve the applicability of techniques like RCA, CCA, and BNA (and related techniques) because, just like standard statistical approaches (like Factor Analysis, Principal Component Analysis, Latent Class Analysis and so forth), and even techniques (like Multiple Correspondence Analysis that are sometimes opposed to the traditional statistical ones, these newfangled relational techniques are also dragged down by the macrogenre curse. Thus, RCA and CCA are bound to compute relationality and correlation distances within people but have to rely on vague macrogenre labels. The resulting evaluation logics are thus interpreted using a macrogenre lens---e.g., ``Anything but Heavy Metal'' or ``Highbrow versus Lowbrow" \citep[e.g.,][]{goldberg2011mapping, willekens2022cultural}. But what if what one group means by ``Heavy Metal" differs from what another group means by the same vague label? Are we certain that certain {\em all} types of classical music are ``opposed'' to or divergent from engagement with Country or Hip Hop? Some versions of Indie/Alt Rock are considered today as ``highbrow'' as Classical music (used to be) \citep{Van_Poecke2018}; or there may exist types of Country music that appeal to highly educated audiences and, thus, to people who might also listen to musicals? All quantitative techniques researchers use in the sociology of taste must work with the (broad) genre categories included in the survey. As such, they are powerless to deal with these (reasonable) objections.

@@

\subsection{Data to be Used in the Study}
The empirical part of the study uses data on the musical tastes of Americans collected with Sara Skiles in the summer of 2012 \citep{lizardo_skiles15, lizardo_skiles16}. The survey was (partially) designed to replicate the General Social Survey (GSS) 1993 Culture Module. This is now a canonical data set providing the empirical basis for a variety of analyses (and reanalyses) in the sociology of taste \citep[e.g.,][]{bryson96, goldberg2011mapping, schultz2010strength, han2003unraveling, tampubolon2008revisiting, okada2017structure}.  The data were collected by an organization then called Survey Sample International (SSI), a private firm specializing in sampling, data collection, and analysis. SSI managed recruitment and participation invitation tasks to generate a panel of adults from which our working sample was drawn. Survey respondents were selected from the panel for participation based on age, gender, race, education, and geographic region to approximate a sample representative of the U.S. population ($n = 2,263$). Like GSS 1993, SSI 2012 included items assessing respondents’ likes and dislikes (as well as a middle category of ``mixed feelings'') for 20 categories of musical style: classical/ symphony and chamber, opera/operetta, jazz, Broadway musicals/show tunes, mood/easy listening, big band/swing, classic rock/oldies, country, bluegrass, folk, hymns/gospel, Latin/Spanish/salsa, rap/hip hop, blues/R\&B, reggae, top 40/pop music, contemporary rock, indie/alternative rock, dance/club/electronic, and hard rock/heavy metal. In addition to providing a taste judgment for each item, respondents were also asked if they regularly listened to each macrogenre. The following analyses are based on a data frame in which people are coded as engaging the macrogenre if they report both liking and listening to it. The two-mode network we will be working with is thus of dimensions and has $2263$ people and $20$ genres. This defines an affiliation matrix $A_{PG}$ of the same dimensions, with entries equal to 1 if person $P$ likes \textit{and} listens to macrogenre label $G$ and zero otherwise.

\subsection{The Link Clustering Approach}
I propose that exploiting {\em relationality} hidden in the usual survey data collecting information on people and genres can help us make headway on the microgenre critique outlined in the preceding. As noted at the outset, this requires that we stop looking at the usual survey data that forms the bulk of empirical material in the quantitative study of taste as ``survey data'' (e.g., data collected on the characteristics of individuals in the form of ``variables''). This is a presumption that is baked into traditional techniques (like Factor Analysis or Latent Class Analysis) that fit statistical models to such data. It is also embedded in ``relational'' techniques like Multiple Correspondence Analysis which, while not steeped in the preoccupations of standard statistical inference (although it could be), divert our attention to the indirect relations between macrogenres (or between people), usually conceptualized as distances in a social space. This kind of indirect relationality is fine and dandy but as already noted in the case of RCA, CCA, and BNA, all the relationality built into MCA is powerless, if what is fed {\em into} the process is the same old, limited, overextended macrogenre categories.

I address the microgenre issue using a technique for overlapping community detection called link clustering \citep{ahn_etal10}. The basic idea is simple and is illustrated in the toy example shown in Figure~\ref{fig:link-toy1}. Breaking with approaches that attempt to cluster people and genres by focusing on the people or the genres, we take the person-to-genre {\em link} as the unit of analysis and focus our classification energies on those. Classifying the links gives (by definition) a classification of the entities at the end of each link (people and genres) for free. Not only that, because most people choose multiple genres, and all genres ``choose'' multiple people, this classification guarantees that genres get assigned to multiple \textit{overlapping} clusters, at the same time splitting up the macrogenres into more relationally focused clumps. This allows us to make headway on both horns of the microgenre critique---namely, overlapping categories and internal differentiation---at once.

\begin{figure}[ht!]
     \centering
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy1.png}
        \caption{Two person-genre edges sharing the same genre.}
        \label{fig:link-toy1}
    \end{subfigure} 
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy2.png}
        \caption{Similarity between person-genre edges depends on the other genres people are linked to.}
        \label{fig:link-toy2}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/toy-sim.png}
        \caption{Edge similarity matrix from toy example.}
        \label{fig:toy-sim}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/toy-dis.png}
        \caption{Edge distance matrix from toy example.}
        \label{fig:toy-dis}
    \end{subfigure}
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy.png}
        \caption{Dendrogram from Ward clustering of toy example edge distance matrix cut at four clusters.}
        \label{fig:link-toy}
    \end{subfigure}
    \caption{Toy illustration of how the link clustering approach works.}
    \label{fig:link-toy-ex}
 \end{figure}

@@

Going back to our running toy example, we are tasked with computing the similarities between $\frac{11 \times (11-1)}{2} = 55$ person to genre pairs. Suppose we wanted to calculate the similarity between the $P_1-G_2$ and $P_3-G_2$ person-to-genre edges in the two-mode network. In Figure~\ref{fig:link-toy1}, these two links are highlighted in yellow the ``strung out'' edgelist and are shown in red in the corresponding bipartite subgraph shown in Figure~\ref{fig:link-toy2}. The similarity ($S$) between the two links is then given by:

@@

Where $n_+(P_1)$ is the cultural ego network of Person 1 and $n_+(P_3)$ is the cultural ego network of Person 3. The formula says that the similarity between two person-to-genre links sharing a genre, in this case, the similarity between the $P_1-G_2$ edge and the $P_3-G_2$ edge, is given by the intersection of Person 1's and Person 3's cultural ego network divided by their union (also known as Jaccard's coefficient). When the cultural ego networks of two people are the same then $S=1$, when they are completely disjoint (share no genres, meaning the intersection is the empty set), then $S = 0$. All other cases of person-to-genre edges sharing the same genre return a number between zero and one ($0 > S < 1$).\footnote{Note that the only substantive similarities we care about are between person-to-genre edges that shared the same genre but have different people attached to them at the other end. All other person-to-genre edges, featuring people connected to different genres are set to zero.} In the toy example case, $P_1$ and $P_3$ consume two genres out of the four possible ones they could consume in common, resulting in an edge similarity score of $\frac{2}{4} = 0.50$.

If we do that for all pairs of person-to-culture genres sharing one genre node in common, we end up with the two-way, one-mode (with the only mode being edges) $11 \times 11$ similarity matrix shown in Figure~\ref{fig:link-toy3}. Note two features of the similarity matrix. First, by definition, two person-to-genre links that go from different genres to the same person are maximally similar (as the overlap is 1.0). This means that only links that go from different people to the same genre exhibit overlap variation, and this is driven (``reflectively'' \citep{lizardo18}) by the overlap between the neighborhoods of the other mode (people). Next, we transform the similarity matrix into a {\em dissimilarity} matrix by subtracting one from each entry (shown in  Figure~\ref{fig:link-toy4}). We then subject this matrix to hierarchical clustering using Ward's \citeyearpar{ward63} method.

The resulting dendrogram from the cluster analysis is shown in Figure~\ref{fig:link-toy5}. Note that moving from the top down, the dendrogram splits person-to-genre links according to the macrogenre label. That is, macrogenre information is preserved in the clustering and can be recovered by splitting the dendrogram at a height where the number of clusters is equal to the original number of macrogenres (in this case, four as shown in Figure~\ref{fig:link-toy5}). Moving from the bottom up, the hierarchical clustering of the dissimilarity matrix yields {\em link communities} \citep{ahn_etal10}, in which both people nodes, but, most crucially, genres nodes are assigned to different clumps. In the limiting case (the dendrogram's bottom-most ``leafs''), each person-to-genre link is assigned to a cluster. The more interesting thing is that as we move up to a point below macrogenres clusters and above the leaves, we see a partition of {\em varieties} of macrogenres. For instance, macrogenre $G4$ is split into two focused microgenre varieties. The first is attached to persons $P2$ and $P4$, and the second attached to persons $P3$ and $P5$, respectively. The idea, is that the macrogenre nodes (e.g., ``Classical,'') assigned to different clumps represent micro-variations of the macrogenre label that differ primarily relationally or structurally; different ``types'' of ``Classical'' are different because their audiences make distinct choices {\em with respect} to the other genres in the survey.

\begin{figure}[ht!]
     \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/all-branches-macro.png}
        \caption{c = 8, k = 20}
        \label{fig:dend-macro}
    \end{subfigure} 
     \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/all-branches-macro-labels.png}
        \caption{c = 8, k = 20}
        \label{fig:dend-macro-labels}
    \end{subfigure} 
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/classic-rock-branches.png}
        \caption{c = 3, k = 12}
        \label{fig:dend-micro-classic-rock}
    \end{subfigure} 
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/blues-branches.png}
        \caption{c = 3, k = 5}
        \label{fig:dend-micro-blues}
    \end{subfigure}
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/opera-branches.png}
        \caption{c = 3, k = 2}
        \label{fig:dend-micro-opera}
    \end{subfigure}
    \caption{}
    \label{fig:dend}
 \end{figure}

@@

\section{Analysis and Results}
\subsection{Discovering microgenre Communities in Real Data}
Let us see how this process works in real data. We will see that link clustering can uncover valid microgenres. Recall that the data features 2,263 people choosing up to 20 macrogenre labels. When strung out as an edgelist, this results in 9,216 person-to-genre connections in the data, the resulting $9216 \times 9,216$ matrix, containing Jaccard similarities among person-to-genre links sharing a node in either of the two modes, is then the input to an agglomerative hierarchical clustering algorithm using Ward's \citep{ward63} method. The hierarchical clustering process proceeds as follows.

Initially, each person-to-genre link is assigned to a singleton cluster. Then, in the second time step, the pairs of links with the largest Jaccard similarities are put in the same clump. This continues at each time step, where pairs of links with the largest similarity are chosen, and their respective communities are merged. This process is repeated until all links belong to a single cluster. The history of the clustering process is then stored in a dendrogram, which encodes all the information on the hierarchical organization of the genre communities. The height of the dendrogram provides information about the strength of the genre communities. As we have seen, the highest levels reproduce the original macrogenres, while the lower levels uncover more focused microgenres embedded within them. This is shown in Figure~\ref{fig:dend}.

Figure~\ref{fig:dend-macro} shows that when we cut the dendrogram at a high level (e.g., $C = 8$), we reproduce the original twenty macrogenres we began with as the ``discovered'' link communities. As Figure~\ref{fig:dend-macro-labels} shows, the agglomerative link clustering procedure roughly arranges the macrogenre levels by their original popularity (number of person-to-genre links). From left to right, these are Classic Rock/Oldies, Pop/Top 40, Country, Classical, Easy Listening, Blues/R\&B, Contemporary Rock, Rap/Hip Hop, Jazz, Gospel, Dance/Club, Indie Alternative, Latin, Broadway/Musicals, Heavy Metal/Hard Rock, Reggae, Big Band, Folk, Bluegrass, and Opera. As shown in Figure~\ref{fig:dend-micro}, microgenre communities are produced by cutting the link clustering dendrogram at a lower level. Choosing a cutoff value of $c = 3$ results in $k = 102$ microgenre communities extracted from the original twenty macrogenre labels. Data scientists wring their hands about choosing a cut value when performing cluster analysis. One advantage of the link clustering approach is that we always know what we are doing because microgenres are strictly nested within the original macrogenres. On the other hand, choosing a smaller cut value produces finer-grained microgenres (perhaps at the expense of analytic tractability and interoperability), and selecting a higher value returns us to broader genre communities closer to the original macrogenre labels we began with (with the limiting case as shown in Figure~\ref{fig:dend-macro} being the original macro genre labels themselves). As Figure~\ref{fig:dend-micro-labels} shows, at any cut value $C$, the more popular macrogenres produce more microgenre communities, while the less popular ones produce a smaller number. For instance, as shown in Figure~\ref{fig:dend-micro-classic-rock}, at $c = 3$, the macrogenre ``Classic Rock/Oldies'' (the leftmost set of branches in the dendrogram in Figure~\ref{fig:dend-macro}) is split (at the point at which the branches intersect the red dashed cut line) into $k_{(CR)} = 12$ distinct micro-variations, while, as Figure~\ref{fig:dend-micro-opera} shows, the macrogenre ``Opera" (the rightmost set of branches in the dendrogram in Figure~\ref{fig:dend-macro}) is split into only $k_{(Op)} = 2$  microgenre communities. Figure~\ref{fig:dend-micro-blues} shows that ``Blues/R\&B'' falls somewhere in between, with $k_{(BRB)} = 5$  microgenre communities uncovered.

Each cut point in the dendrogram will have a microgenre size distribution $N$ associated with it. Lower cut points return a microgenre size distribution dominated by many microgenres with tiny ($N_{mg} < 10$) or trivial $N_{mg} = 2$ audiences. In this respect, although I have been using the labels ``macro'' and ``micro'' as if they referred to substantive or objective partitions, they are best interpreted as {\em relative} to a given classification level. Thus, microgenres are ``micro'' relative to the usual (perhaps ``basic'' in Rosch's \citeyearpar{Rosch1978-ue} sense) categorization level of vague macrogenre labels populating most arts participation surveys. But these may be micro, as we will see, relative to the (logics, discourses, schemas) meta-groupings that we generate when applying the usual suite of statistical and data reduction techniques (e.g., such as ``highbrow'' or ``Folk"). Further, at any given classification level, further microgenres will be nested within any given micro-partition. In this way, microgenre communities reflect a substantively valid way in which people partition the cultural world, whereby there is always the possibility of making finer-grained distinctions within fine-grained distinctions (e.g., ``old school rap," ``1980s old school rap,", ``early 1980s old school rap," and the like).

>>>>

\subsection{Precursors of the Proposed Approach}
The approach proposed here builds on the core ideas of ``duality'' and ``relationality'' that have become foundational for formal approaches to measuring culture in the sociology of taste \citep{mutzel2020duality, mohr2015formal}. In this sense, I followed a path already trailed by others. For instance, Goldberg's \citeyearpar{goldberg2011mapping} {\em Relational Class Analysis} (RCA) ``classifies the people'' (in the terms developed below when reviewing standard data-analytic approaches in the sociology of taste) included a survey by partitioning a person-by-person matrix built from higher order similarities (sums of differences of differences) between each pair of person's corresponding response vector coding their stances (e.g., like, dislike, neutral) toward a set of cultural objects (e.g., genres, attitude items).

Boutyline's 
\citeyearpar{boutyline2017improving} {\em Correlational Class Analysis} (CCA) also classifies the people but this time by using the absolute value of the correlation distance between response vectors to build the person-by-person similarity matrix. Boutyline and Vaisey's {\em Belief Network Analysis} (BNA) classifies the objects people connect to (in their case, beliefs, but could also be genres) by creating a weighted object network, where the weight of the tie between two objects (attitude toward government spending and attitude toward immigration) is given by the magnitude of their correlation coefficient in the survey.

As we will see, I propose a complementary and synergistic approach with these previous efforts. More to the point, the relational identification of microgenres should be able to improve the applicability of techniques like RCA, CCA, and BNA (and related techniques) because, just like standard statistical approaches (like Factor Analysis, Principal Component Analysis, Latent Class Analysis and so forth), and even techniques (like Multiple Correspondence Analysis that are sometimes opposed to the traditional statistical ones, these newfangled relational techniques are also dragged down by the macrogenre curse. Thus, RCA and CCA are bound to compute relationality and correlation distances within people but have to rely on vague macrogenre labels.

The resulting evaluation logics are thus interpreted using a macrogenre lens---e.g., ``Anything but Heavy Metal'' or ``Highbrow versus Lowbrow" \citep[e.g.,][]{goldberg2011mapping, willekens2022cultural}. But what if what one group means by ``Heavy Metal" differs from what another group means by the same vague label? Are we certain that certain {\em all} types of classical music are ``opposed'' to or divergent from engagement with Country or Hip Hop? Some versions of Indie/Alt Rock are considered today as ``highbrow'' as Classical music (used to be) \citep{Van_Poecke2018}; or there may exist types of Country music that appeal to highly educated audiences and, thus, to people who might also listen to musicals? All quantitative techniques researchers use in the sociology of taste must work with the (broad) genre categories included in the survey. As such, they are powerless to deal with these (reasonable) objections.

@@

\section{Analytic Appproach}
\subsection{Data to be Used in the Study}
The empirical part of the study uses data on the musical tastes of Americans collected with Sara Skiles in the summer of 2012 \citep{lizardo_skiles15, lizardo_skiles16}. The survey was (partially) designed to replicate the General Social Survey (GSS) 1993 Culture Module. This is now a canonical data set providing the empirical basis for a variety of analyses (and reanalyses) in the sociology of taste \citep[e.g.,][]{bryson96, goldberg2011mapping, schultz2010strength, han2003unraveling, tampubolon2008revisiting, okada2017structure}.  The data were collected by an organization then called Survey Sample International (SSI), a private firm specializing in sampling, data collection, and analysis. SSI managed recruitment and participation invitation tasks to generate a panel of adults from which our working sample was drawn. Survey respondents were selected from the panel for participation based on age, gender, race, education, and geographic region to approximate a sample representative of the U.S. population ($n = 2,263$).

Like GSS 1993, SSI 2012 included items assessing respondents’ likes and dislikes (as well as a middle category of ``mixed feelings'') for 20 categories of musical style: classical/ symphony and chamber, opera/operetta, jazz, Broadway musicals/show tunes, mood/easy listening, big band/swing, classic rock/oldies, country, bluegrass, folk, hymns/gospel, Latin/Spanish/salsa, rap/hip hop, blues/R\&B, reggae, top 40/pop music, contemporary rock, indie/alternative rock, dance/club/electronic, and hard rock/heavy metal. In addition to providing a taste judgment for each item, respondents were also asked if they regularly listened to each macrogenre. The following analyses are based on a data frame in which people are coded as engaging the macrogenre if they report both liking and listening to it. The two-mode network we will be working with is thus of dimensions and has $2263$ people and $20$ genres. This defines an affiliation matrix $A_{PG}$ of the same dimensions, with entries equal to 1 if person $P$ likes \textit{and} listens to macrogenre label $G$ and zero otherwise.

\subsection{The Link Clustering Approach}
I propose that exploiting {\em relationality} hidden in the usual survey data collecting information on people and genres can help us make headway on the microgenre critique outlined in the preceding. As noted at the outset, this requires that we stop looking at the usual survey data that forms the bulk of empirical material in the quantitative study of taste as ``survey data'' (e.g., data collected on the characteristics of individuals in the form of ``variables''). This presumption is baked into traditional techniques (like Factor Analysis or Latent Class Analysis) that fit statistical models to such data. It is also embedded in ``relational'' techniques like Multiple Correspondence Analysis which, while not steeped in the preoccupations of standard statistical inference (although it could be), divert our attention to the indirect relations between macrogenres (or between people), usually conceptualized as distances in a social space. This kind of indirect relationality is fine and dandy. Still, as already noted in the case of RCA, CCA, and BNA, all the relationality built into MCA is powerless if what is fed {\em into} the process is the same old, limited, overextended macrogenre categories.

\subsubsection{A Toy Example}
I address the microgenre issue using a technique for overlapping community detection called \textit{link clustering} \citep{ahn_etal10}. The basic idea is simple and illustrated in the toy example in Figure~\ref{fig:link-toy1}. Breaking with approaches that attempt to cluster people and genres by focusing on the people or the genres, we take the person-to-genre {\em link} as the unit of analysis and focus our classification energies on those. Classifying the links gives (by definition) a classification of the entities at the end of each link (people and genres) for free. Moreover, because most people choose multiple genres, and all genres ``choose'' multiple people, this classification guarantees that genres get assigned to multiple \textit{overlapping} clusters while splitting up the macrogenres into more relationally focused clumps. This allows us to make headway on both horns of the microgenre critique---namely, overlapping categories and internal differentiation---at once.

\begin{figure}[t!]
     \centering
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy1.png}
        \caption{Two person-genre edges sharing the same genre.}
        \label{fig:link-toy1}
    \end{subfigure} 
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy2.png}
        \caption{Similarity between person-genre edges depends on the other genres people are linked to.}
        \label{fig:link-toy2}
    \end{subfigure}
\end{figure}

@@

Going back to our running toy example, we are tasked with computing the similarities between $\frac{11 \times (11-1)}{2} = 55$ person-to-genre pairs. Suppose we wanted to calculate the similarity between the $P_1-G_2$ and $P_3-G_2$ person-to-genre edges in the two-mode network. In Figure~\ref{fig:link-toy1}, these two links are highlighted in yellow by the ``strung out'' edgelist and are shown in red in the corresponding bipartite subgraph shown in Figure~\ref{fig:link-toy2}. The similarity ($S$) between the two links is then given by:

@@

Where $n_+(P_1)$ is the cultural ego network of Person 1 and $n_+(P_3)$ is the cultural ego network of Person 3. The formula says that the similarity between two person-to-genre links sharing a genre, in this case, the similarity between the $P_1-G_2$ edge and the $P_3-G_2$ edge, is given by the intersection of Person 1's and Person 3's cultural ego network divided by their union (also known as Jaccard's coefficient). When the cultural ego networks of two people are the same $S=1$, when they are completely disjoint, sharing no genres, meaning the intersection is the empty set,  $S = 0$. All other cases of person-to-genre edges sharing the same genre return a number between zero and one ($0 > S < 1$).\footnote{Note that the only substantive similarities we care about are between person-to-genre edges that share the same genre but have different people attached to them at the other end. All other person-to-genre edges featuring people connected to different genres are set to zero.} In the toy example case, $P_1$ and $P_3$ consume two genres out of the four possible ones they could consume in common, resulting in an edge similarity score of $\frac{2}{4} = 0.50$.

\begin{figure}
    \centering
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/toy-sim.png}
        \caption{Edge similarity matrix from toy example.}
        \label{tab:toy-sim}
    \end{subfigure}
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/toy-dis.png}
        \caption{Edge distance matrix from toy example.}
        \label{tab:toy-dis}
    \end{subfigure}
     \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{Toy/link-clust-toy.png}
        \caption{Dendrogram from Ward clustering of toy example edge distance matrix cut at four clusters.}
        \label{fig:link-toy}
    \end{subfigure}
    \caption{Toy illustration of how the link clustering approach works.}
    \label{fig:link-toy-ex}
 \end{figure}

If we do that for all pairs of person-to-culture genres sharing one genre node in common, we end up with the two-way, one-mode (with the only mode being the edges) $11 \times 11$ similarity matrix shown in Figure~\ref{tab:toy-sim}. Note two features of the similarity matrix. First, by definition, two person-to-genre links that go from different genres to the same person are maximally similar (as the overlap is 1.0). This means that only links that go from different people to the same genre exhibit overlap variation, and this is driven (``reflectively'' \citep{lizardo18}) by the overlap between the neighborhoods of the other mode (people). Next, we transform the similarity matrix into a {\em dissimilarity} matrix by subtracting one from each entry (shown in  Figure~\ref{tab:toy-dis}). We then subject this matrix to hierarchical clustering using Ward's \citeyearpar{ward63} method.

The resulting dendrogram from the cluster analysis is shown in Figure~\ref{fig:link-toy}. Note that moving from the top down, the dendrogram splits person-to-genre links according to the macrogenre label. That is macrogenre information is preserved in the clustering and can be recovered by splitting the dendrogram at a height where the number of clusters is equal to the original number of macrogenres (in this toy case, four). Moving from the bottom up, the hierarchical clustering of the dissimilarity matrix yields {\em link communities} \citep{ahn_etal10}, in which both people nodes, but, most crucially, genres nodes are assigned to different clumps. In the limiting case, each person-to-genre link is assigned to a cluster, namely, the dendrogram's bottom-most ``leaves.'' The more interesting thing is that as we move up to a point below macrogenres clusters and above the leaves, we see a partition of {\em varieties} of macrogenres. For instance, macrogenre $G4$ is split into two focused microgenre varieties. The first is attached to persons $P2$ and $P4$, and the second is attached to persons $P3$ and $P5$, respectively. The idea is that the macrogenre nodes (e.g., ``Classical'') assigned to different clumps represent micro-variations of the macrogenre label that differ primarily relationally or structurally; different ``types'' of ``Classical'' are different because their audiences make distinct choices {\em with respect} to the other genres in the survey.

\begin{figure}[t!]
     \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/all-branches-macro.png}
        \caption{c = 8, k = 20}
        \label{fig:dend-macro}
    \end{subfigure} 
     \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/all-branches-macro-labels.png}
        \caption{c = 8, k = 20}
        \label{fig:dend-macro-labels}
    \end{subfigure} 
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/classic-rock-branches.png}
        \caption{c = 3, k = 12}
        \label{fig:dend-micro-classic-rock}
    \end{subfigure} 
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/blues-branches.png}
        \caption{c = 3, k = 5}
        \label{fig:dend-micro-blues}
    \end{subfigure}
     \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{Plots/Dend/opera-branches.png}
        \caption{c = 3, k = 2}
        \label{fig:dend-micro-opera}
    \end{subfigure}
    \caption{}
    \label{fig:dend}
 \end{figure}

@@

\subsection{Discovering Microgenre Communities in Real Data}
Let us see how this process works in real data. Recall that the data features 2,263 people choosing up to 20 macrogenre labels. When strung out as an edgelist, this results in 9,216 person-to-genre connections in the data, the resulting $9216 \times 9,216$ matrix, containing Jaccard similarities among person-to-genre links sharing a node in either of the two modes, is then the input to an agglomerative hierarchical clustering algorithm using Ward's \citep{ward63} method. The hierarchical clustering process proceeds as follows. Initially, each person-to-genre link is assigned to a singleton cluster. Then, in the second time step, the pairs of links with the largest Jaccard similarities are put in the same clump. This continues at each time step, where pairs of links with the largest similarity are chosen, and their respective communities are merged. This process is repeated until all links belong to a single cluster. The history of the clustering process is then stored in a dendrogram, which encodes all the information on the hierarchical organization of the genre communities. The height of the dendrogram provides information about the strength of the genre communities. As we have seen, the highest levels reproduce the original macrogenres, while the lower levels uncover more focused microgenres embedded within them. This is shown in Figure~\ref{fig:dend}.

Figure~\ref{fig:dend-macro} shows that when we cut the dendrogram at a high level (e.g., $C = 8$), we reproduce the original twenty macrogenres we began with as the ``discovered'' link communities. As Figure~\ref{fig:dend-macro-labels} shows, the agglomerative link clustering procedure roughly arranges the macrogenre levels by their original popularity (number of person-to-genre links). From left to right, these are Classic Rock/Oldies, Pop/Top 40, Country, Classical, Easy Listening, Blues/R\&B, Contemporary Rock, Rap/Hip Hop, Jazz, Gospel, Dance/Club, Indie Alternative, Latin, Broadway/Musicals, Heavy Metal/Hard Rock, Reggae, Big Band, Folk, Bluegrass, and Opera. As shown in Figures~\ref{fig:dend-micro-classic-rock}-\ref{fig:dend-micro-opera}, microgenre communities for specific genres are produced by cutting the link clustering dendrogram at a lower level; in the examples provided ($C = 3$).

Data scientists worry constantly about choosing a cut value when performing cluster analysis. One advantage of the link clustering approach is that we always know what we are doing because microgenres are strictly nested within the original macrogenres. Choosing a smaller cut value produces finer-grained microgenres (perhaps at the expense of analytic tractability and interoperability), and selecting a higher value returns us to broader genre communities closer to the original macrogenre labels we began with (with the limiting case as shown in Figure~\ref{fig:dend-macro} being the original macro genre labels themselves). At any cut value $C$, the more popular macrogenres produce more microgenre communities, while the less popular ones produce a smaller number.

For instance, as shown in Figure~\ref{fig:dend-micro-classic-rock}, at $c = 3$, the macrogenre ``Classic Rock'' (the leftmost set of branches in the dendrogram in Figure~\ref{fig:dend-macro}) is split (at the point at which the branches intersect the red dashed cut line) into $k_{(CR)} = 12$ distinct micro-variations, while, as Figure~\ref{fig:dend-micro-opera} shows, the macrogenre ``Opera" (the rightmost set of branches in the dendrogram in Figure~\ref{fig:dend-macro}) is split into only $k_{(Op)} = 2$  microgenre communities. Figure~\ref{fig:dend-micro-blues} shows that ``Blues/R\&B'' falls somewhere in between, with $k_{(BRB)} = 5$  microgenre communities uncovered. Each cut point in the dendrogram will have a microgenre size distribution $N$ associated with it. Lower cut points return a microgenre size distribution dominated by many microgenres with tiny ($N_{mg} < 10$) or trivial $N_{mg} = 2$ audiences. Another approach, and the one I will be using in the case studies that follow, is to pick a macrogenre label (a branch in the larger dendrogram), decide \textit{ex ante} how many microgenres it makes sense to analyze, and then cut that branch of the dendrogram at a value that returns the desired number of microgenres.

Note, that although I have been using the labels ``macro'' and ``micro'' as if they referred to substantive or objective partitions, they are best interpreted as {\em relative} to a given classification level. Thus, microgenres are ``micro'' relative to the usual (perhaps ``basic'' in Rosch's \citeyearpar{Rosch1978-ue} sense) categorization level of vague macrogenre labels populating most arts participation surveys. At any given classification level, further microgenres will be nested within any given micro-partition. In this way, microgenre communities should reflect a substantively valid way in which people partition the cultural world, whereby there is always the possibility of making finer-grained distinctions within fine-grained distinctions (e.g., ``old school rap," ``1980s old school rap," ``early 1980s old school rap," and the like).

\section{Two Genre Case Studies}
We have shown that we can perform a link clustering and that this clustering returns a nice set of microgenres nested within the usual macrogenres in the survey. But does this exercise result in any substantive gains? In this section, I show that the link clustering approach does allow you to extract new insights from old data. There are many ways we can proceed. I choose to present two ``case studies'' of macrogenre labels and the resulting microgenre variations returned by the link clustering procedure. My criteria were as follows: I selected two macrogenre labels for which I could find some substantiated claim in the scholarly literature regarding possible microgenre variation. Naturally, general claims about microgenre variation apply to all macrogenre labels. Therefore, the ``substantiated in the scholarly literature" serves as a way to narrow the field. I also intended to identify macrogenre labels for which some sort of debate or uncertainty existed in the scholarly literature regarding their overall status and their variations. I settled on Heavy Metal and Latin/Spanish Music/Salsa using this criterion. As we will see, Metal definitely fits this bill, as a lively debate already exists regarding its status as non-elite or an elite cultural form \citep{tampubolon2008revisiting}. Latin/Spanish/Salsa represents a more ``minor'' case study, but one that has also been documented to exhibit the sort of microgenre variation we seek \citep{bakk2013estimating}. As such, Salsa also showcases the usefulness of the link clustering approach for resolving key issues in the literature and extracting new insights from existing data.

\subsection{Heavy Metal}
Heavy Metal is one of the most storied macrogenre labels in the sociology of taste, as featured in the title of Bethany Bryson's \citeyearpar{bryson96} now classic paper. Since then, it has been the subject of scholarly debate regarding its possibly changing status in the cultural stratification order \citep{tampubolon2008revisiting, goldberg2011mapping, lizardo_skiles15}. Heavy Metal stood out in Bryson's study for several reasons. First, it was the most disliked genre in the 1993 General Social Survey culture module data---one of the most heavily analyzed data sources in the sociology of taste---hence the ``Anything But'' in the title of Bryson's eponymous piece. Second, it was the prototypical genre that was relatively more favored by working-class audiences but absolutely hated by individuals with high education. Thus, it served as the linchpin of Bryson's symbolic exclusion argument; omnivores were omnivorous up to a point, except when it came to those genres most favored by those with low education.
